---
title: "Reproduce paper results"
output: 
  html_document:
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Model overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE)
knitr::opts_knit$set(root.dir = "~/InfectionTrees")
devtools::load_all()
library(InfectionTrees)
library(tidyr)
library(dplyr)
library(kableExtra)
library(ggplot2)
theme_set(theme_bw() + theme(axis.title = element_text()))
```

## Preliminaries

1. Make sure `InfectionTrees` is installed (`devtools::install_github("skgallagher/InfectionTrees")`) and loaded along with the following libraries

```{r eval = FALSE}
library(InfectionTrees)
library(ggplot2)
library(knitr)
library(kableExtra)
library(tidyr)
library(dplyr)
```

2. If there are any occurrences of `devtools::load_all()`, remove them and replace with `library(InfectionTrees)`.

3.  Expected run time is an approximation for of estimated time to run the script.  They were originally run on a laptop with the following specs unless otherwise specified.

```{r eval = FALSE}
Distributor ID:	Ubuntu
Description:	Ubuntu 20.04.2 LTS
Release:	20.04
Codename:	focal

Architecture:                    x86_64
CPU op-mode(s):                  32-bit, 64-bit
Byte Order:                      Little Endian
Address sizes:                   39 bits physical, 48 bits virtual
CPU(s):                          8
On-line CPU(s) list:             0-7
Thread(s) per core:              2
Core(s) per socket:              4
Socket(s):                       1
NUMA node(s):                    1
Vendor ID:                       GenuineIntel
CPU family:                      6
Model:                           142
Model name:                      Intel(R) Core(TM) i7-10510U CPU @ 1.80GHz
Stepping:                        12
CPU MHz:                         1281.004
CPU max MHz:                     4900.0000
CPU min MHz:                     400.0000
BogoMIPS:                        4599.93
Virtualization:                  VT-x
L1d cache:                       128 KiB
L1i cache:                       128 KiB
L2 cache:                        1 MiB
L3 cache:                        8 MiB
NUMA node0 CPU(s):               0-7
Vulnerability Itlb multihit:     KVM: Mitigation: VMX disabled
Vulnerability L1tf:              Not affected
Vulnerability Mds:               Not affected
Vulnerability Meltdown:          Not affected
Vulnerability Spec store bypass: Mitigation; Speculative Store Bypass disabled v
                                 ia prctl and seccomp
Vulnerability Spectre v1:        Mitigation; usercopy/swapgs barriers and __user
                                  pointer sanitization
Vulnerability Spectre v2:        Mitigation; Enhanced IBRS, IBPB conditional, RS
                                 B filling
Vulnerability Srbds:             Mitigation; TSX disabled
Vulnerability Tsx async abort:   Not affected
Flags:                           fpu vme de pse tsc msr pae mce cx8 apic sep mtr
                                 r pge mca cmov pat pse36 clflush dts acpi mmx f
                                 xsr sse sse2 ss ht tm pbe syscall nx pdpe1gb rd
                                 tscp lm constant_tsc art arch_perfmon pebs bts 
                                 rep_good nopl xtopology nonstop_tsc cpuid aperf
                                 mperf pni pclmulqdq dtes64 monitor ds_cpl vmx e
                                 st tm2 ssse3 sdbg fma cx16 xtpr pdcm pcid sse4_
                                 1 sse4_2 x2apic movbe popcnt tsc_deadline_timer
                                  aes xsave avx f16c rdrand lahf_lm abm 3dnowpre
                                 fetch cpuid_fault epb invpcid_single ssbd ibrs 
                                 ibpb stibp ibrs_enhanced tpr_shadow vnmi flexpr
                                 iority ept vpid ept_ad fsgsbase tsc_adjust bmi1
                                  avx2 smep bmi2 erms invpcid mpx rdseed adx sma
                                 p clflushopt intel_pt xsaveopt xsavec xgetbv1 x
                                 saves dtherm ida arat pln pts hwp hwp_notify hw
                                 p_act_window hwp_epp md_clear flush_l1d arch_ca
                                 pabilities

```

## File Structure

All code for reproducing our revised paper results may be found [here](https://github.com/skgallagher/InfectionTrees/tree/master/revised_paper_results).

Please clone the [entire repository](https://github.com/skgallagher/InfectionTrees/) into your local directory and set your working directory to `revised_paper_results/`

The results include the following:

1. Table 1 results. `table1-binary-cov-base-model-sims/`
    a. Base model simulation results. `run-base-sims.R`
    b. Base model simulation bootstrap results. `run-base-sims-boot.R`
    c. Table output for base model simulation. `analyze-base-sims.R`
2. Table 2 results. `table2-binary-cov-mot-model-sims/`
    a. Multiple outside transmissions model simulation results. `outsider-sims.R`
    c. Table output for base model simulation. `analyze-mot-sims.R`
3. The TB data results. `md-tb/`
    a. Tables 3 and 4. `tab3-4-data-eda.R`
    b. Tables 5 and 6. 
        i. Results from base and multiple outside transmissions model. `tb-data-base-and-mot.R`
        ii. Bootstrap standard error for base and multiple outside transmissions model. `bootstrap-sims.R`
        iii. Naive model results. `data-naive-model.R`
        iv. Naive model bootstrap error. `data-naive-boot.R`
        v. Table outputs for the three models combined. `analyze-data-big-table.R`
    c. Figure 4, most likely trees for the base model. `fig4-most-likely-trees.R`



## Table 1 results. `table1-binary-cov-base-model-sims/`

This table was produced by running the following three files in order.  

###  Base model simulation results. `run-base-sims.R`

```{r echo = FALSE}
tb1_fn <- "~/InfectionTrees/revised_paper_results/table1-binary-cov-base-model-sims/run-base-sims.R"


```


Expected run-time: ~3-5 hours

```{r, code = readLines(tb1_fn), eval = FALSE, results = 'hide'}
```


###  Base model bootstrap simulation results. `run-base-sims-boot.R`

```{r echo = FALSE}
tb1_fn <- "~/InfectionTrees/revised_paper_results/table1-binary-cov-base-model-sims/run-base-sims-boot.R"


```


Expected run-time: ~12-18 hours

```{r, code = readLines(tb1_fn), eval = FALSE, results = 'hide'}
```

### Table output for base model simulation. `analyze-base-sims.R`

```{r echo = FALSE}
tb1_fn <- "~/InfectionTrees/revised_paper_results/table1-binary-cov-base-model-sims/analyze-base-sims.R"


```


Expected run-time: ~seconds, given the above completed results.

```{r, code = readLines(tb1_fn), eval = FALSE, results = 'hide'}
```


## Table 2 results. `table2-binary-cov-mot-model-sims/`

This table was produced by running the following two files in order.  

###  Multiple outside transmissions model simulation results. `outsider-sims.R`

```{r echo = FALSE}
tb1_fn <- "~/InfectionTrees/revised_paper_results/table1-binary-cov-base-model-sims/run-base-sims.R"


```


Expected run-time: ~30 hours

```{r, code = readLines(tb1_fn), eval = FALSE, results = 'hide'}
```



### Table output for multiple outside transmissions model simulation. `analyze-mot-sims.R`

```{r echo = FALSE}
tb1_fn <- "~/InfectionTrees/revised_paper_results/table2-binary-cov-mot-model-sims/analyze-mot-sims.R"


```


Expected run-time: ~seconds, given the above completed results.

```{r, code = readLines(tb1_fn), eval = FALSE, results = 'hide'}
```

##  The TB data results. `md-tb/`

### Tables 3 and 4. `tab3-4-data-eda.R`

```{r echo = FALSE}
tb1_fn <- "~/InfectionTrees/revised_paper_results/md-tb/tab3-4-data-eda.R"


```


Expected run-time: ~seconds

```{r, code = readLines(tb1_fn), eval = FALSE, results = 'hide'}
```
