---
title: "TB Data Results: Visualization of clusters"
date: "`r Sys.Date()`"
output:
    html_document:
      code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
                        cache = TRUE,
                        warning = FALSE,
                        message = FALSE
)
devtools::load_all()

library(tidyr)
library(dplyr)
library(kableExtra)
library(ggplot2)
library(forcats)
library(readr)
theme_set(theme_bw() + theme(axis.title = element_text()))
```

## Libraries

```{r eval = FALSE}
library(InfectionTrees)
library(tidyr)
library(dplyr)
library(kableExtra)
library(ggplot2)
library(forcats)
library(readr)
```


## Vignette goals

1. Recap our [model fitting results from the TB data](tb-data-base.html)

2. We will visualize the best model in the data by
    a. Looking at within-cluster transmission
    b. Look at the most likely clusters by generation

3. Sample new clusters from our best model

## Recapping our best model

We have run the code (but do not show) that was used to fit our base model to the TB data.  For more details see the [data results section](tb-data-base.html).



We can approximate the probability that individual an individuals with probability $p_A$ of infecting an individual with probability $p_B$.  We provide the function `mc_trees_to_A_infects_B` which calculates this probability for you for a given cluster.  We look at cluster 27.  For this exercise, we sample a new set of MC trees than what was used to fit the model.

```{r echo = FALSE}
data(tb_clean)
    clusters <- tb_clean %>%
        dplyr::mutate(smear = ifelse(spsmear == "Positive",
                                     1, 0),
                      cluster_id = group,
                      hiv_f = hivstatus) %>%
        dplyr::mutate(hiv_neg_pos = ifelse(hiv_f == "neg", 1, 0),
               hiv_unk_pos = ifelse(hiv_f == "unk", 1, 0)) %>%
        dplyr::group_by(cluster_id) %>%
        dplyr::mutate(rel_time = rel_time / 365) %>%
        dplyr::mutate(cluster_size = dplyr::n()) %>%
        dplyr::ungroup() %>%
        dplyr::mutate(race_f = forcats::fct_collapse(race,
                                     white = "White",
                                     black = "Black or African American",
                                     asian = "Asian")) %>%
        dplyr::mutate(race_asian_white = ifelse(race_f == "asian", 1, 0),
               race_black_white = ifelse(race_f == "black", 1, 0)) %>%
        dplyr::select(cluster_id, smear,
               hiv_neg_pos,
               hiv_unk_pos,
               rel_time,
               race_asian_white,
               race_black_white,
               cluster_size)

```

```{r}

par_ests <- c("Intercept" = -.72690792,
              "smear" = -.09322607,
              "hiv_neg_pos" = -0.36415841,
              "hiv_unk_pos" = -0.56817810,
              "rel_time" = .345552628)



mc_trees <-  sample_mc_trees(clusters %>% filter(cluster_id == 27),
                                 B = 10000,
                                 multiple_outside_transmissions = FALSE,
                                 covariate_names = names(par_ests))
```


```{r}
plotting_df <- mc_trees_to_A_infects_B(my_id = 27,
                                   par_ests = par_ests,
                                   mc_trees = mc_trees,
                                   output = "tidy")


```


We can then plot the associations, for example, with a heat map.  The output of `mc_trees_to_A_infects_B()` is a long data frame of probability that $A$ infected $B$, which is estimated over the MC samples.



```{r}
ggplot(data = plotting_df) +
    geom_tile(aes(y = factor(infector_id),
                  x = factor(infected_id),
                  fill = prob * 100)) +
    theme_bw(base_size = 10) +
    labs(title = "Estimated transmission probability of TB from person A to B",
         subtitle = sprintf("Individuals in cluster %s", 27),
         y = "Infector ID",
         x = "Infected ID") +
    theme(legend.position = "bottom",
          panel.grid.major = element_blank(),
          panel.border = element_blank()) +
     scale_fill_gradient(low = "gray90",
                        high = "black",
                        name = latex2exp::TeX("Prob. A infects B (%)"),
                        na.value = "white") 

```
