---
title: "TB Data Results: Visualization of clusters"
date: "`r Sys.Date()`"
output:
    html_document:
      code_folding: hide
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
                        cache = TRUE,
                        warning = FALSE,
                        message = FALSE
)
devtools::load_all()

library(tidyr)
library(dplyr)
library(kableExtra)
library(ggplot2)
library(forcats)
library(readr)
theme_set(theme_bw() + theme(axis.title = element_text()))
```

## Libraries

```{r eval = FALSE}
library(InfectionTrees)
library(tidyr)
library(dplyr)
library(kableExtra)
library(ggplot2)
library(forcats)
library(readr)
```


## Vignette goals

1. Recap our [model fitting results from the TB data](tb-data-base.html)

2. We will visualize the best model in the data by
    a. Looking at within-cluster transmission
    b. Look at the most likely clusters by generation

3. Sample new clusters from our best model

## Recapping our best model

We have run the code (but do not show) that was used to fit our base model to the TB data.  For more details see the [data results section](tb-data-base.html).

```{r echo = FALSE, results = 'hide'}
clusters <- tb_clean %>%
    filter(PCR.Cluster != "") %>%
    mutate(smear = ifelse(spsmear == "Positive",
                          1, 0),
           cluster_id = PCR.Cluster,
           hiv_f = fct_collapse(hivstatus,
                              pos = c("Positive"),
                              neg = c("Negative"),
                              unk = c("Unknown", "Not offered",
                                      "Refused")),
           age = parse_number(ageatrept)) %>%
    mutate(hiv_neg_pos = ifelse(hiv_f == "neg", 1, 0),
           hiv_unk_pos = ifelse(hiv_f == "unk", 1, 0)) %>%
    group_by(cluster_id) %>%
    mutate(rel_time = as.numeric(datecoll - min(datecoll)) / 365) %>%
    mutate(cluster_size = n()) %>%
    ungroup() %>%
    mutate(race_f = fct_collapse(race,
                                 white = "White",
                                 black = "Black or African American",
                                 asian = "Asian")) %>%
    mutate(race_asian_white = ifelse(race_f == "asian", 1, 0),
           race_black_white = ifelse(race_f == "black", 1, 0)) %>%
    select(cluster_id, smear,
           hiv_neg_pos,
           hiv_unk_pos,
           rel_time,
           race_asian_white,
           race_black_white,
           cluster_size)

```

```{r echo = FALSE, results = 'hide'}
K <- 1000
my_seed <- 42
set.seed(my_seed)

## MODELS
## models
covariate_list <- vector(mode = "list", length = 5)
covariate_list[[1]] <- NA
covariate_list[[2]] <- "smear"
covariate_list[[3]] <- c("smear",
                         "hiv_neg_pos", "hiv_unk_pos")
covariate_list[[4]] <- c("smear",
                         "hiv_neg_pos", "hiv_unk_pos",
                         "rel_time")
covariate_list[[5]] <- c("smear",
                         "hiv_neg_pos", "hiv_unk_pos",
                         "rel_time",
                         "race_asian_white",
                         "race_black_white")

## Set up outputs
loglike_df <- data.frame(model = 1:length(covariate_list),
                         n_params = c(1, 2, 4, 5, 7),
                         loglike = 0,
                         aic = 0)
beta_mat1 <- matrix(0, nrow = 1, ncol = 4)
rownames(beta_mat1) <- c("Intercept")
colnames(beta_mat1) <- c("Est.", "lower", "upper", "SE")
beta_list <- vector(mode = "list", length = length(covariate_list))
beta_list[[1]] <- beta_mat1
for(ii in 2:length(covariate_list)){
  mat <- matrix(0, nrow = length(covariate_list[[ii]]) + 1,
                ncol = 4)
  rownames(mat) <- c("Intercept", covariate_list[[ii]])
  colnames(mat) <- c("Est.", "lower", "upper", "SE")
  beta_list[[ii]] <- mat
}

```

```{r echo = FALSE, results = 'hide'}
t_init <- proc.time()[3]

## Sample MC trees all at once

    t0 <- proc.time()[3]
    ## Sample all MC clusters at once for each of the 5 models
    mc_trees <-  sample_mc_trees(clusters,
                                 B = K,
                                 multiple_outside_transmissions = FALSE,
                                 covariate_names = covariate_list[[length(covariate_list)]])
    print(proc.time()[3] - t0)


## Fit each of the models

for(jj in 1:length(covariate_list)){
        covariate_names <- covariate_list[[jj]]
        print("Model:")
        print(covariate_names)
        if(is.na(covariate_names[1])){
            init_params <- 0
        } else{
            init_params <- rep(0, length(covariate_names) + 1)
        }
        ## Optimize

        print("Optimizing")
        bds <- rep(-5, length(init_params))
        if(length(covariate_names) > 5){
            bds <- rep(-4, length(init_params))
        }
        lower_bds <- bds
        upper_bds <- -bds
        cov_mat <- covariate_df_to_mat(mc_trees,
                                       cov_names = covariate_names)

    t1 <- proc.time()[3]
    best_params <- optim(par = init_params,
                         fn = general_loglike,
                         mc_trees = data.table::as.data.table(mc_trees),
                         return_neg = TRUE,
                         cov_mat = cov_mat,
                         cov_names = covariate_names,
                         use_outsider_prob = FALSE,
                         multiple_outside_transmissions = FALSE,
                         method = "L-BFGS-B",
                         lower = lower_bds,
                         upper = upper_bds,
                         hessian = TRUE
                         )
    t2 <- proc.time()[3] - t1
    print(paste("Optimization time:", round( t2 / 60, 3),
                "min"))

    beta_list[[jj]][, 4] <- sqrt(diag(solve(best_params$hessian))) ## SE from Fisher info

    ## Likelihood Profile ests
    print("Likelihood profiling")
    lp_ests <- likelihood_profs(best_pars = best_params$par,
                                    max_loglike = -best_params$value,
                                    mc_trees = mc_trees,
                                    covariate_names = covariate_names,
                                    alpha = .05,
                                    multiple_outside_transmissions = FALSE)

    beta_list[[jj]][,-4] <- lp_ests

    print("best params:")
    print(beta_list[[jj]])
    loglike_df$loglike[jj] <- -best_params$val

    print(paste("Total time:", round( (proc.time()[3] - t_init)  / 3600, 3),
                "hrs"))


}
```

The best base model has the following covariates and coefficient estimates:

```{r}
beta_list[[4]] %>%
  kable(digits = 2) %>%
  kable_styling(bootstrap_options = c("condensed", "hover", "striped", "responsive"),
                full_width = FALSE, position = "center")
```

