---
title: "First look at the Tuberculosis datas"
output: html_document
vignette: >
  %\VignetteIndexEntry{Model overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE)
devtools::load_all()
library(InfectionTrees)
library(tidyr)
library(dplyr)
library(kableExtra)
library(ggplot2)
theme_set(theme_bw() + theme(axis.title = element_text()))
```


##  Overview

<<<<<<< HEAD
In this vignette, we will explore the data set provided here in `data(tb_clean)`, which is a data set originally collected and analyzed in [Xie et al. (2018)](https://academic.oup.com/cid/article-abstract/67/11/1653/4984625)
=======
In this vignette, we will explore the data set provided here in `data(tb_clean)`, which is a data set originally collected and analyzed in [Xie et al. (2018)](GETLINK-and-REF)
>>>>>>> fd400a1555a7a782265dd95dc3e8cb9f965adce0


## Exploratory Data Analysis


```{r, eval = FALSE, echo = TRUE}
library(InfectionTrees)
library(tidyverse)
library(kableExtra)

```

<<<<<<< HEAD
The data `tb_clean` has `r nrow(tb_clean)` rows and `r ncol(tb_clean)` columns, where each row in the data corresponds to an individual with a detected case of Tuberculosis (TB).

The columns include demographic characteristics about the patients as well as two diagnostic tests of the disease, including a AFB sputum smear test, which we analyze here.


```{r}
tb_sub <- tb_clean %>%
  mutate(cluster_id = group) %>%
=======
The data `tb_clean` has `r nrow(tb_clean)` rows and `r ncol(tb_clean)` columns, where each row in the data corresponds to an individual who has been diagnosed with Tuberculosis (TB).

The columns include demographic characteristics about the patients as well as two diagnostic tests of the disease, including a AFB sputum smear test, which we analyze here.

Of the `r nrow(tb_clean)` cases, `r sum(tb_clean$PCR.Cluster != "")` indivduals have been clustered together through PCR of the TB culture.

```{r}
tb_sub <- tb_clean %>%
  filter(PCR.Cluster != "") %>%
  mutate(cluster_id = PCR.Cluster) %>%
>>>>>>> fd400a1555a7a782265dd95dc3e8cb9f965adce0
  group_by(cluster_id) %>%
        mutate(cluster_size = n()) %>%
  ungroup()

```


<<<<<<< HEAD
This data has `r length(unique(tb_sub$cluster_id))` clusters with the following distribution of cluster sizes:
=======
This subset of data has `r length(unique(tb_sub$cluster_id))` clusters with the following distribution of cluster sizes:
>>>>>>> fd400a1555a7a782265dd95dc3e8cb9f965adce0

```{r}
sizes <- tb_sub %>% group_by(cluster_id) %>%
  summarize(cluster_size = n()) %>% 
  ungroup() %>%
  group_by(cluster_size) %>%
  summarize(freq = n()) 
  
  
sizes %>%
  kable() %>%
  kable_styling(bootstrap_options = c("condensed", "hover", "striped", "responsive"),
                full_width = FALSE, position = "center")

                    
```


This means that `r round(sizes$freq[sizes$cluster_size == 1] / sum(sizes$freq) * 100, 1)`% of the clusters are singletons and `r round(sum(sizes$freq[sizes$cluster_size %in% 1:3]) / sum(sizes$freq) * 100, 1)`% of the clusters have size at most 3.

A single cluster of individuals looks like the following:

```{r}
<<<<<<< HEAD
tb_sub %>% filter(group == "27") %>%
  arrange(rel_time) %>%
  select(group, rel_time, everything()) %>%
=======
tb_sub %>% filter(PCR.Cluster == "27") %>%
  arrange(datecoll) %>%
  select(cluster_id, datecoll, everything()) %>%
>>>>>>> fd400a1555a7a782265dd95dc3e8cb9f965adce0
   kable() %>%
  kable_styling(bootstrap_options = c("condensed", "hover", "striped", "responsive"),
                full_width = FALSE, position = "center")
```


In our data set of `r nrow(tb_sub)` individuals, `r sum(tb_sub$hivstatus == "Positive")` or `r round(sum(tb_sub$hivstatus == "Positive") / nrow(tb_sub) * 100)`% of individuals are HIV+.  In the clusters with more than 3 individuals, `r sum(tb_sub$hivstatus == "Positive" & tb_sub$cluster_size > 3)` of `r sum(tb_sub$cluster_size > 3)` or `r round(sum(tb_sub$hivstatus == "Positive" & tb_sub$cluster_size > 3) / sum(tb_sub$cluster_size > 3) * 100)`% of individuals are HIV+. 
