---
title: "First look at the Tuberculosis data"
output: html_document
vignette: >
  %\VignetteIndexEntry{Model overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      cache = TRUE,
                      warning = FALSE,
                      message = FALSE)
devtools::load_all()
library(InfectionTrees)
library(tidyr)
library(dplyr)
library(kableExtra)
library(ggplot2)
theme_set(theme_bw() + theme(axis.title = element_text()))
```


##  Overview


In this vignette, we will explore the data set provided here in `data(tb_clean)`, which is a data set originally collected and analyzed in [Xie et al. (2018)](https://academic.oup.com/cid/article-abstract/67/11/1653/4984625).


## Exploratory Data Analysis


```{r, eval = FALSE, echo = TRUE}
library(InfectionTrees)
library(tidyverse)
library(kableExtra)

```


The data `tb_clean` has `r nrow(tb_clean)` rows and `r ncol(tb_clean)` columns, where each row in the data corresponds to an individual with a detected case of Tuberculosis (TB).

The columns include demographic characteristics about the patients as well as two diagnostic tests of the disease, including a AFB sputum smear test, which we analyze here.


```{r}
tb_sub <- tb_clean %>%
  mutate(cluster_id = group) %>%
  group_by(cluster_id) %>%
        mutate(cluster_size = n()) %>%
  ungroup()

```



This data has `r length(unique(tb_sub$cluster_id))` clusters with the following distribution of cluster sizes:


```{r}
sizes <- tb_sub %>% group_by(cluster_id) %>%
  summarize(cluster_size = n()) %>% 
  ungroup() %>%
  group_by(cluster_size) %>%
  summarize(freq = n()) 
  
  
sizes %>%
  kable() %>%
  kable_styling(bootstrap_options = c("condensed", "hover", "striped", "responsive"),
                full_width = FALSE, position = "center")

                    
```


This means that `r round(sizes$freq[sizes$cluster_size == 1] / sum(sizes$freq) * 100, 1)`% of the clusters are singletons and `r round(sum(sizes$freq[sizes$cluster_size %in% 1:3]) / sum(sizes$freq) * 100, 1)`% of the clusters have size at most 3.

A single cluster of individuals looks like the following:

```{r}
tb_sub %>% filter(group == "27") %>%
  arrange(rel_time) %>%
  select(group, rel_time, everything()) %>%
   kable() %>%
  kable_styling(bootstrap_options = c("condensed", "hover", "striped", "responsive"),
                full_width = FALSE, position = "center")
```


In our data set of `r nrow(tb_sub)` individuals, `r sum(tb_sub$hivstatus == "Positive")` or `r round(sum(tb_sub$hivstatus == "Positive") / nrow(tb_sub) * 100)`% of individuals are HIV+.  In the clusters with more than 3 individuals, `r sum(tb_sub$hivstatus == "Positive" & tb_sub$cluster_size > 3)` of `r sum(tb_sub$cluster_size > 3)` or `r round(sum(tb_sub$hivstatus == "Positive" & tb_sub$cluster_size > 3) / sum(tb_sub$cluster_size > 3) * 100)`% of individuals are HIV+. 
